---
output: html_document
---

# Big Title



## Little Title



### Even Littler Title


Normal text, just perfectly normal text, nothing special here.
If we want to jazz it up a bit we can use *italics* for example, or even **bold** text!

Making a list of other things we can do:  

*  write equations such as $$ e = m c^2 $$
*  add url links: [click here](http://www.google.com)
*  do numbered lists - that are indented further
    1. like
    2. this
    
And tables are quite straightforward as well:


First Header  | Second Header | Third Header
------------- | ------------- | ------------
Content Cell  | Content Cell  |  More stuff



## OK, now for some data analysis

First let's get some data, one of the builtin datasets will do, iris sepal lentgh is a favourite.

From the documentation: 
> Fisher's or Anderson's iris data set gives the measurements in centimeters of the variables sepal length and width and petal length and width, respectively, for 50 flowers from each of 3 species of iris. The species are *Iris setosa*, *versicolor*, and *virginica*.

Here is a quick look at the data set, to give you a flavour: 

```{r}
head(iris)

```

Just quickly, there are is `r nrow(iris)` cases or rows in the dataset, and the mean petal lenght is 
`r mean(iris$Petal.Length)`. 


### Simple scatterplot

Now we can explore the data visually a bit, like plotting the petal length by its width:

```{r}
plot(iris$Petal.Length, iris$Petal.Width, 
     xlab=c("Petal lenght"),
     ylab=c("Petal width"),
     main="Iris Data")

```

### Linear regression

How about a linear regression line in the chart? And colour coding the data points by species?
First let's use a linear model  and see what the intercept and slope are?

```{r}
lm(iris$Petal.Width ~ iris$Petal.Length )$coefficients
```

OK, not we can do the same plot as before, but add colours and the line. And because this is pretty straightforward, I won't bore you with the code, but only show you the plot instead:

```{r, echo=FALSE}
plot(iris$Petal.Length, iris$Petal.Width, 
     col=c("red","green","blue")[unclass(iris$Species)],
     xlab=c("Petal lenght"),
     ylab=c("Petal width"),
     main="Iris Data")
abline(lm(iris$Petal.Width ~ iris$Petal.Length )$coefficients, col="black")

```

### More to come...