---
output: html_document
runtime: shiny
---

# Big Title
  
  
  
## Little Title
  
  
  
### Even Littler Title
  

Normal text, just perfectly normal text, nothing special here.
If we want to jazz it up a bit we can use *italics* for example, or even **bold** text!
  
  Making a list of other things we can do:  
  
*  write equations such as $$ e = m c^2 $$
*  add url links: [click here](http://www.google.com)
*  do numbered lists - that are indented further
  1. like
  2. this

And tables are quite straightforward as well:
  
  
  First Header  | Second Header | Third Header
------------- | ------------- | ------------
  Content Cell  | Content Cell  |  More stuff



## OK, now for some data analysis

First let's get some data, one of the builtin datasets will do, iris sepal lentgh is a favourite.

From the documentation: 
> Fisher's or Anderson's iris data set gives the measurements in centimeters of the variables sepal length and width and petal length and width, respectively, for 50 flowers from each of 3 species of iris. The species are *Iris setosa*, *versicolor*, and *virginica*.

Here is a quick look at the data set, to give you a flavour: 

```{r}
head(iris)

```

Just quickly, there are is `r nrow(iris)` cases or rows in the dataset, and the mean petal lenght is 
`r mean(iris$Petal.Length)`. 


### Simple scatterplot

Now we can explore the data visually a bit, like plotting the petal length by its width:
```{r}
plot(iris$Petal.Length, iris$Petal.Width, 
xlab=c("Petal lenght"),
ylab=c("Petal width"),
main="Iris Data")

```

### Linear regression

How about a linear regression line in the chart? And colour coding the data points by species?
First let's use a linear model  and see what the intercept and slope are?

```{r}
lm(iris$Petal.Width ~ iris$Petal.Length )$coefficients
```

OK, not we can do the same plot as before, but add colours and the line. And because this is pretty straightforward, I won't bore you with the code, but only show you the plot instead:

```{r, echo=FALSE}
plot(iris$Petal.Length, iris$Petal.Width, 
col=c("red","green","blue")[unclass(iris$Species)],
xlab=c("Petal lenght"),
ylab=c("Petal width"),
main="Iris Data")
abline(lm(iris$Petal.Width ~ iris$Petal.Length )$coefficients, col="black")

```


### More stuff - by Joe Cheng <joe@rstudio.com>


```{r, echo = FALSE}
shinyApp(

ui = pageWithSidebar(
headerPanel('Iris k-means clustering'),
sidebarPanel(
selectInput('xcol', 'X Variable', names(iris)),
selectInput('ycol', 'Y Variable', names(iris),
selected=names(iris)[[2]]),
numericInput('clusters', 'Cluster count', 3,
min = 1, max = 9)
),
mainPanel(
plotOutput('plot1')
)
),

palette(c("#E41A1C", "#377EB8", "#4DAF4A", "#984EA3",
"#FF7F00", "#FFFF33", "#A65628", "#F781BF", "#999999")),

server = function(input, output, session) {

# Combine the selected variables into a new data frame
selectedData <- reactive({
iris[, c(input$xcol, input$ycol)]
})

clusters <- reactive({
kmeans(selectedData(), input$clusters)
})

output$plot1 <- renderPlot({
par(mar = c(5.1, 4.1, 0, 1))
plot(selectedData(),
col = clusters()$cluster,
pch = 20, cex = 3)
points(clusters()$centers, pch = 4, cex = 4, lwd = 4)
})

})

```